### 2021-07-21 x1_covid.Rmd

---
title: "COVID"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

covidregionaldata := one of the many packages dealing with pre-packaging COVID data.
Potential sources WHO (often cited as unreliable), ECDC (European Center of Disease Control - changed recently to weekly data - provides R tailored download), and JHU (kicked off data collection/harmonisation effort, often cited/used)

```{r, message=FALSE}
library(covidregionaldata)
library(dplyr)
library(tidyr)
library(ggplot2)
library(scales)
```

```{r}
covid_jhu <- get_national_data(source = "jhu")
covid_who <- get_national_data(source = "who")
covid_ecdc<- get_national_data(source = "ecdc")
```

```{r}
global_jhu <- covid_jhu %>%
  group_by(date) %>%
  summarise( cases_total     = sum(cases_total, na.rm = TRUE)
            ,deaths_total    = sum(deaths_total, na.rm = TRUE)
            ,recovered_total = sum(recovered_total, na.rm = TRUE))
```

```{r}
global_jhu %>% 
  pivot_longer(cols = 2:4, names_to = "category", values_to = "counts") %>%

ggplot(aes(x = date, y = counts, fill = category)) +
  geom_area() +
  labs(title = "JHU ???")
```

```{r}
global_who <- covid_who %>%
  group_by(date) %>%
  summarise( cases_total     = sum(cases_total, na.rm = TRUE)
            ,deaths_total    = sum(deaths_total, na.rm = TRUE)
            ,recovered_total = sum(recovered_total, na.rm = TRUE))
```

```{r}
global_who %>% 
  pivot_longer(cols = 2:4, names_to = "category", values_to = "counts") %>%

ggplot(aes(x = date, y = counts, fill = category)) +
  geom_area() +
  labs(title = "WHO - no recovered cases")
```

```{r}
global_ecdc <- covid_ecdc %>%
  group_by(date) %>%
  summarise( cases_total     = sum(cases_total, na.rm = TRUE)
            ,deaths_total    = sum(deaths_total, na.rm = TRUE)
            ,recovered_total = sum(recovered_total, na.rm = TRUE))
```

```{r}
global_ecdc %>% 
  pivot_longer(cols = 2:4, names_to = "category", values_to = "counts") %>%

ggplot(aes(x = date, y = counts, fill = category)) +
  geom_area() +
  labs(title = "ECDC")
```

test another COVID19 project

```{r}
#install.packages("coronavirus")
# library(coronavirus)
# ---------- it is better to pull in latest changes by calling without library 
coronavirus::update_dataset()  # pulls latest (data)version
library(coronavirus)           # then call library

covid <- coronavirus %>% as_tibble()
```

```{r}
global_cov <- covid %>% 
  group_by(type, date) %>%
  summarise(total_cases = sum(cases), .groups = "drop") %>%
  pivot_wider(names_from = type, values_from = total_cases) %>%
  arrange(date) %>%
  mutate(active = confirmed - death - recovered) %>%
  mutate(active_total = cumsum(active),
                recovered_total = cumsum(recovered),
                death_total = cumsum(death))

global_cov
```

```{r}
tmp <- global_cov %>%
  select(date, contains("total")) %>%
  pivot_longer(cols = 2:4, names_to = "category", values_to = "counts")

tmp %>%

ggplot(aes(x = date, y = counts, fill = category)) +
  geom_area() +
  theme_minimal() +
  labs(x = NULL, y = NULL) +
  theme(legend.position = c(0.2, 0.3)) +
  scale_fill_manual( name = "cases"
                    ,values = c("orange","red", "green")
                    ,labels = c("active", "death","recovered")) +
  scale_y_continuous(labels = scales::label_number_si())
```

```{r}
conf_df <- covid %>% 
  filter(type == "confirmed") %>%
  group_by(country) %>%
  summarise(total_cases = sum(cases)) %>%
  arrange(-total_cases) %>%
  mutate(
     share  = round( 100 * (total_cases / sum(total_cases)))
    ,parents = "Confirmed cases") %>%
  ungroup() 
```

```{r}
# plotly based treemap
library(plotly)
 plot_ly(data = conf_df,
          type= "treemap",
          values = ~total_cases,
          labels= ~ country,
          parents=  ~parents,
          domain = list(column=0),
          name = "Confirmed",
          textinfo="label+value+percent parent")
```

```{r}
# install.packages("treemapify")
# for options and how to use check out: https://r-charts.com/part-whole/treemapify/

library(treemapify)

ggplot( data = conf_df
       , aes( area  = total_cases, fill = as.factor(share), colour = I("white")
             ,label = paste0(country, "\n", total_cases, "\n", share, "%")
             )
       ) +
  geom_treemap(start = "topleft") +
  geom_treemap_text(colour = "white", start = "topleft") +
  theme(legend.position = "none") +
  #scale_fill_brewer(palette = "Blues")
  scale_fill_viridis_d()
```