### 2021-07-22 paper-references.Rmd
arrrggghh ... working on references in word mode :(

---
title: "Assessing the Global COVID19 Impact on Air Transport with Open Data"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
# failed to render and I could not finally track what kills the pdf
 bookdown::pdf_book:
   base_format: rticles::ieee_article
  # bookdown::word_document2:
  # #   reference_docx: tmpl_Conference-template-letter.docx
  #   default
bibliography: "DASC2021.bib"
---

This helper document serves exclusively to get out the references from Zotero and in a paste and copy formatting.

EUROCONTROL [@statfor_growth_2018], for example,

Reported number of cases (22. Jan 2020 through 27. May 2021) [@krispin_coronavirus_2021] based on data from John Hopkins University [@JHU_2020]

global daily flight data set [@xavier_olive_2021_4893103].

RStudio [@rstudio] tools for the R language [@R]

15 minutes or more [@strohmeier_crowdsourced_2021].

PRR 2020 [@prr_2020]

who declaring covid [@who_covid_2020]

BBC 2020 [@bbc_coronavirus_2020]

ABC 2020 australia [@worthington_covid_aus_2020]

France24 corona-americas [@france24_coronavirus_2020]

ICAO [@icao_covid_impact_2021]

resilience [@resilience_engineering_2009]

glu [@gluchshenko_2012]

D1.3 [@gluchshenko_d1_res2050_2013]

koelle [@koelle_resilience_2015]

REFERENCES ============================