### 2021-07-24 paper-visualisations.Rmd
submitted version (Word & pdf), some supporting files for the final push

---
title: "Untitled"
author: "PRU"
date: "23/07/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(roll)
```

Final push analysis

- data geo-referenced to increase hits --> base_osn_2_

```{r, message=FALSE}
## load study data - as long as not outsourced to data prep
fns <- list.files(path = "./data/", pattern = "base_osn_2_", full.names = TRUE )
col_spec <- cols(
  ADEP = col_character(),
  ADEP_CTRY = col_character(),
  ADEP_REG = col_character(),
  ADES = col_character(),
  ADES_CTRY = col_character(),
  ADES_REG = col_character(),
  TYPE = col_character(),
  DATE = col_date(format = ""),
  CALL = col_character()
)
ds  <- fns %>% map_dfr(.f = ~ read_csv(.x))
```

load aircraft data base
```{r}
ac_db <- read_csv("./data-raw/aircraftDatabase.csv"
                  ,col_types = cols(.default = "c"))

#ac_types <- read_csv("./data-raw/doc8643AircraftTypes.csv")

ds <- ds %>% left_join(ac_db %>% select(icao24, registration, typecode, icaoaircrafttype), by = c("ICAO24"="icao24"))

```

## getting the meta data table

```{r}
daily_tfc <- read_csv("./data/daily_osn.csv", col_types = cols()) %>%
  #--------- correct for double counts in January 2019
  group_by(DATE) %>%
  mutate(COUNT = row_number()) %>%
  filter(COUNT == 1)

monthly_tfc <- daily_tfc %>% mutate(MOF = lubridate::floor_date(DATE, unit = "month")) %>%
  group_by(MOF) %>% summarise(N = n(), FLTS = sum(FLTS))

dwnlds <-

table_1 <- data.frame(
   FNS  = fns
  ,SIZE = file.info(fns)$size     # this is the size of our base_osn files - redo!
  ,FLTS = monthly_tfc$FLTS 
  ,YEAR = c(rep(2019, 12), rep(2020, 12), rep(2021, 6))
  )

paper_tbl <- table_1 %>%
  group_by(YEAR) %>%
  summarise(SIZE = sum(SIZE), FLTS = sum(FLTS))
```

##

```{r}
# summarise regional level
## -------- percentile thresholds
pcts   <- c(0.8, 0.6, 0.3, 0.1)

## -------- function definition
regional_timeline <- function(.ds, .reg, .pcts=pcts){
us <- .ds %>% 
  filter(ADEP_REG == .reg | ADES_REG == .reg) %>%
  group_by(YR = year(DATE), DATE) %>% 
  summarize(N = n(), .groups = "drop")

# static thresholds, x-th percentile rounded to lower integer
us_ref <- us %>% filter(YR == 2019) %>%
  pull(N) %>% quantile(probs = .pcts) %>% floor()

us_ref2 <- us %>% filter(YR == 2019, DATE >= lubridate::ymd("2019-07-01")) %>%
  pull(N) %>% quantile(probs = .pcts) %>% floor()

# use % of base level
base <- us_ref2[1]    # 80th percentile
cut_lvls <- base * c(1, 0.8, 0.6, 0.4, 0.2)
cut_lvls <- floor(cut_lvls)

max_N <- max(us$N)
lvls  <- c("Nom", "Base", "-20%", "-40%", "-60%","-80","low")
us <- us %>% 
  mutate( #THRES  = cut(N, breaks = c(max_N, us_ref, 0),  labels = lvls)
         #,
         THRES2 = cut(N, breaks = c(max_N, cut_lvls, 0)
                      #, labels = lvls
                      ,dig.lab=12    # to force full numbers
                      )
         )
return(list(us, us_ref, us_ref2, cut_lvls))
}

## do for us
res <- regional_timeline(ds, "US")
us  <- res[[1]]
us_ref2 <- res[[3]]
cut_lvls <- res[[4]]
```

```{r fig-us-timeline, fig.cap="(ref:fig-us-timeline)"}
plot_timeline <- function(.df, .pct_ref){
 p <- .df %>%
  ggplot(aes(x = DATE)) +
  geom_hline(yintercept = .pct_ref, linetype = "dotdash") +
  geom_point(aes(y = N, color = THRES2), alpha = 0.3) +
  geom_line( aes(y = N, colour = THRES2, group = 1)) +
  theme_minimal()
 return(p)
}

 gg <- plot_timeline(us, cut_lvls)

 gg 
 #+
#   coord_cartesian(ylim=c(0, 7))
```
```{r}
us_rq <- ds %>% filter(ADEP_REG == "US" | ADES_REG == "US")
us_rq %>% summarise( N = n()
                    , NOT_P = sum(!icaoaircrafttype %in% c("L1P","L2P"))
                    , NOT_PH= sum(!icaoaircrafttype %in% c("L1P","L2P","H1P", "H1T","H2T"))
)

us_rq <- us_rq %>% filter(!icaoaircrafttype %in% c("L1P","L2P","H1P", "H1T","H2T"))                                  
```

```{r}
res <- regional_timeline(us_rq, "US")
us  <- res[[1]]
cut_lvls <- res[[4]]

lbl_df <- data.frame(DATE = lubridate::ymd("2019-07-1")
                     ,LVLS= cut_lvls) %>%
  mutate(LBL = paste0(c("base", "-20%","-40%","-60%","-80%")," (", LVLS, ")")  )

gg <- plot_timeline( us %>% filter(DATE >= lubridate::ymd("2019-07-01"))
                    ,cut_lvls) 
gg +
  geom_text(lbl_df, mapping = aes(x = DATE, y = LVLS, label = LBL)
            , nudge_y = 1500, nudge_x = 30) +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none")
```

```{r}
ggsave("./figures/us-resilience-timeline.png", dpi = 320, height = 6, width = 8)
```

```{r}
check_offset  <- function(.tfc, .ref){
  df <- .tfc %>% 
    mutate(DIFF = N - .ref
           ,CUM_DIFF = if_else(DIFF >= 0, 0, DIFF)
           ,MOF = lubridate::floor_date(DATE, unit = "month"))
  return(df)
}
resilience_dt <- function(.tfc, .ref){
  df <- .tfc %>%
    check_offset(.ref) %>%
    group_by(MOF) %>%
    summarise( N_DISR = sum(DIFF < 0)
              ,FIRST  = DATE[which(DIFF < 0)[1]]
              ,CUM_DIFF = sum(CUM_DIFF))
  return(df)
}
```

```{r}
resi_us <- resilience_dt (
   us %>% filter(DATE >= lubridate::ymd("2019-07-01"))
  ,cut_lvls[2] 
)
resi
```

```{r}
bra <- ds %>% filter(ADEP_REG == "BR" | ADES_REG == "BR") %>%
  filter(!icaoaircrafttype %in% c("L1P","L2P","H1P", "H1T","H2T"))

plot_me <- function(.tfc, .reg, .nudge_y){
res <- regional_timeline(.tfc, .reg)
us  <- res[[1]]
cut_lvls <- res[[4]]

lbl_df <- data.frame(DATE = lubridate::ymd("2019-07-1")
                     ,LVLS= cut_lvls) %>%
  mutate(LBL = paste0(c("base", "-20%","-40%","-60%","-80%")," (", LVLS, ")")  )

gg_bra <- plot_timeline( us %>% filter(DATE >= lubridate::ymd("2019-07-01"))
                    ,cut_lvls) 
gg_bra <- gg_bra +
  geom_text(lbl_df, mapping = aes(x = DATE, y = LVLS, label = LBL)
            , nudge_y = .nudge_y, nudge_x = 30) +
  labs(x = NULL, y = NULL) +
  theme(legend.position = "none")
return(gg_bra)
}
```

```{r}
( g_bra <- plot_me(bra, "BR", 75) )
# ggsave("./figures/bra-resilience-timeline.png", dpi = 320, height = 6, width = 8)
```

```{r}
eur <- ds %>% filter(ADEP_REG == "EU" | ADES_REG == "EU") %>%
  filter(!icaoaircrafttype %in% c("L1P","L2P","H1P", "H1T","H2T"))

( g_eur <- plot_me(eur, "EU", 1000) )
# ggsave("./figures/eur-resilience-timeline.png", dpi = 320, height = 6, width = 8)
```

```{r, message=FALSE}
library(patchwork)
g_eur + g_bra
```

```{r}
resi_us <- resi

bra_res <- regional_timeline(bra, "BR")
resi_bra <- resilience_dt(bra_res[[1]], bra_res[[4]][2]) %>%
  filter(MOF >= lubridate::ymd("2019-07-01"))

eur_res <- regional_timeline(eur, "EU")
resi_eur <- resilience_dt(eur_res[[1]], eur_res[[4]][2]) %>%
  filter(MOF >= lubridate::ymd("2019-07-01"))
```

show this in the presentation

```{r}
aggregate_resilience <- function(.ds){
  my_agg <- .ds %>% 
    mutate(YR = lubridate::year(MOF)) %>% 
    filter(YR %in% c(2020, 2021)) %>% 
    group_by(YR) %>% 
    summarise(N_DISR = sum(N_DISR), sum(CUM_DIFF))
}

us_res <- aggregate_resilience(resi_us) %>% mutate(REG = "US")
eur_res<- aggregate_resilience(resi_eur) %>% mutate(REG= "EU")
bra_res<- aggregate_resilience(resi_bra) %>% mutate(REG= "BR")

res_all <- bind_rows(us_res, eur_res, bra_res)
res_all %>% select(REG, everything())
```

## Airport level analysis

```{r}
apt_timeline <- function(.df){
  df <- .df %>% group_by(DATE) %>%
    summarise(N_DEP = n()
              ,DEP_DOM = sum(ADEP_CTRY == ADES_CTRY)
              ,DEP_REG = sum(ADEP_REG == ADES_REG))
}

plot_timeline <- function(.df){
  df <- .df %>% 
  mutate(N_DEP_SM = roll_quantile(N_DEP, width = 7, p = .50))

  p <- df %>%  
  ggplot() +
  geom_line(aes(x = DATE, y = N_DEP), colour = "grey60") +
  geom_line(aes(x = DATE, y = N_DEP_SM)) +
  scale_x_date( limits = c(as.Date("2020-01-01"), as.Date("2021-06-01"))
               ,date_breaks = "60 days", date_labels = "%b\n%y") +
  labs(x = NULL, y = NULL) +
  theme_minimal()

  return(p)
}

dep_vs_domshare <- function(.df){
  p <- .df %>%
  ggplot() +
  geom_point(aes(y = N_DEP, x = DEP_REG / N_DEP, colour = as.factor(year(DATE)))) +
  labs(colour = NULL, x = NULL, y = NULL
       ,subtitle = "share regional departures") + 
  scale_x_continuous(labels = scales::percent_format(accuracy = 1)
                     ,expand = c(0,0), limits = c(0,NA)) +
  theme_minimal() +
  theme(legend.position = c(0.1, 0.8))
return(p)
}

check_apt <- function(.mvts, .apt){
  apt_mvts <- .mvts %>%
    filter(ADEP == .apt) %>% 
    apt_timeline() %>%
    filter(N_DEP > 10)   # remove outlier

  p1 <- apt_mvts  %>% plot_timeline()
  p2 <- apt_mvts %>% dep_vs_domshare()

  p <- (p1 + labs(subtitle = .apt)) + p2
return(p)
}
```

```{r, message=FALSE}
apt  <- "EGLL"
mvts <- ds %>% filter(ADEP == apt) 
check_apt(mvts, apt)
ggsave(paste0("./figures/", apt, "-combi.png"), dpi = 320, height = 4, width = 8)
```

ggsave("./figures/egll-combi.png", dpi = 320, height = 4, width = 8)

```{r, message=FALSE}
apt  <- "KATL"
mvts <- ds %>% filter(ADEP == apt) 
check_apt(mvts, apt)
ggsave(paste0("./figures/", apt, "-combi.png"), dpi = 320, height = 4, width = 8)
```

```{r, message=FALSE}
apt  <- "SBGR"
mvts <- ds %>% filter(ADEP == apt) 
check_apt(mvts, apt)
ggsave(paste0("./figures/", apt, "-combi.png"), dpi = 320, height = 4, width = 8)
```

## COVID vs national level

```{r}
coronavirus::update_dataset()  # pulls latest (data)version
library(coronavirus)           # then call library

covid <- coronavirus %>% as_tibble()
```

```{r}
ctry <- "BR"

deps <- bra %>% filter(ADEP_CTRY == ctry)
deps <- deps %>% group_by(DATE, ADEP_REG, ADES_REG) %>% summarise(N_DEP = n())
```

```{r}
deps %>% ggplot() + geom_line(aes(x = DATE, y = N_DEP, group = ADES_REG, colour = ADES_REG))
 covs <- covid %>% filter(country == "Brazil") %>% filter(type == "death") %>% mutate(cum_cases = cumsum(cases)) %>% ggplot() + geom_line(aes(x = date, y = cum_cases))
```

```{r}
deps_covs <- deps %>% 
  left_join( covid %>% filter(country == "Brazil", type == "death") %>% 
              mutate(cum_cases = cumsum(cases)) %>% 
               select(DATE = date, CUM_DEATH = cum_cases)
             , by = "DATE"
             )
```

```{r}
coeff <- 500

ggplot(data = deps_covs %>% filter(DATE >= lubridate::ymd("2020-01-01"))
       , aes(x = DATE)) +
  geom_line(aes( y = N_DEP, colour = ADES_REG)) +
  geom_line(aes( y = CUM_DEATH / coeff)) +
 scale_y_continuous(
#  scale_y_log10(
    # Features of the first axis
    name = "departures",

    # Add a second axis and specify its features
    sec.axis = sec_axis( trans= ~ .*coeff, name="confirmed deaths", labels = scales::comma)
  ) +
  geom_vline(xintercept = lubridate::ymd("2020-03-15"), linetype = "dotdash") +
  labs(x = NULL, colour = NULL) +
  theme_minimal() +
  theme(legend.position = c(0.4,0.7))
```

ggsave("./figures/bra-dep-death.png", dpi = 320, width = 8, height = 5)

```{r}
deps <- eur %>% filter(ADEP_CTRY == "DE") %>% 
  group_by(DATE, ADEP_REG, ADES_REG) %>% summarise(N_DEP = n(), .groups = "drop")

deps_covs <- deps %>% 
  left_join( covid %>% filter(country == "Germany", type == "death") %>% 
              mutate(cum_cases = cumsum(cases)) %>% 
               select(DATE = date, CUM = cum_cases)
             , by = "DATE"
             )
```

```{r}
coeff <- 35

ggplot(data = deps_covs %>% filter(DATE >= lubridate::ymd("2020-01-01"))
       , aes(x = DATE)) +
  geom_line(aes( y = N_DEP, colour = ADES_REG)) +
  geom_line(aes( y = CUM / coeff)) +
 scale_y_continuous(
#  scale_y_log10(
    # Features of the first axis
    name = "departures",

    # Add a second axis and specify its features
    sec.axis = sec_axis( trans= ~ .*coeff, name="confirmed deaths", labels = scales::comma)
  ) +
  geom_vline(xintercept = lubridate::ymd("2020-03-15"), linetype = "dotdash") +
  labs(x = NULL, colour = NULL) +
  theme_minimal() +
  theme(legend.position = c(0.3,0.8))
```

ggsave("./figure/ger-dep-death.png", dpi = 320, width = 8, height = 5)